---
title: "Home"
output: html_document
---


## R Markdown

Introduction
This report focuses on the incarceration of the black population in the United States. Mass incarceration refers to the imprisonment of large number of people in government correctional facilities.The US incarcerates more people than any other nation on the globe. This report takes into consideration the Alabama state marked as AL in the dataset columns. 

Variables Considered
This report takes into consideration three variables from the dataset documentation cookbook. These variables are;
- black_prison_pop: Represents the total number of black inmates in the Federal prisons in  different years according to the states and counties with the states.
- white_prison_pop: Represents the total number of white inmates in the Federal prisons in  different years according to the states and counties with the states.
- total_prison_adm_rate: The rate of admission for new prisoners in the Federal Correctional Facilities.



Relevant values of interest (black_prison_pop for Alabama State between 1990-2015)
Median - 85
Mean - 233.6
Maximum - 3917.0
Minimum - 0

Note that these values of interest are limited to the Alabama State only.

Question 1: Trends Over Time Chart

```{r pressure, echo=FALSE}

#################################question 1 : Trends Over Time Chart 

# Load ggplot2
library("ggplot2")

#load data from the csv file ans store in a variable for global access
trends <- read.csv("incarceration_trends.csv")

#print the data to console
trends

#all the data was loaded to the trends variables. 
#we extract the data we want only and continue to plotting the graph
#here, we use the subset function to extract the data we from the dataset
data <- subset(trends, state== "AL" & year>1990)

head(data)

plot(data$year, data$black_prison_pop, type= 'l', 
     ylab='Black Prison Population', xlab = 'Years',main = 'Population of Incarcerated Black People in AL State from 1990 - 2015')



```


Question 2: Black  vs Whit Prisoners Population Over Time in Years

```{r pressure, echo=FALSE}
 ####################### Question 2: Variable Comparison Chart###########

#load tidyverse
library(tidyverse)


#load data from the csv file ans store in a variable for global access
trends <- read.csv("incarceration_trends.csv")

#print the data to console
trends

#here, we use the subset function to extract the data we from the dataset
data <- subset(trends, state== "AL" & year>1990)

#now we extract the two data variables that we need to compare
mainData<- data.frame(year= data$year,
                      black_prisoners = data$black_prison_pop,
                      white_prisoners = data$white_prison_pop,
                      prison_admission = data$total_prison_adm_rate
)
#set the new data frame as the head
head(mainData)
# Line plot in ggplot2 with groups
mainData %>% gather(key = "Legends", value = "Prison_Population", -year) %>%
  ggplot(aes(x = year, y = Prison_Population, color = Legends), ggtitle("Black Vs White Incarceration")) +
  geom_smooth(method = 'lm') +
  scale_color_manual(values = c("tomato3","goldenrod","skyblue2"))





```

## Including Plots

Question 3: Geographic Distribution Map

```{r pressure, echo=FALSE}
#################################################question 3 : Geographic Distribution dataMap #########
#load the maps library 
library(usmap)
# Load ggplot2
library("ggplot2")

#load data from the csv file ans store in a variable for global access
trends <- read.csv("incarceration_trends.csv")

#print the data to console
trends


#set the data to display in the map
data <- subset(trends, year>2005)

#Now display the data on the Map
plot_usmap(
  data = data, values = "black_prison_pop" , color = "red"
) + 
  scale_fill_continuous(
    low = "white", high = "red", name = "Number of Incarcerated People", label = scales::comma
  ) + 
  labs(title = "Incarcelation of Black People Across the United States of America", subtitle = "Date Range : 2005-2015") +
  theme(legend.position = "right")


```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
